---
output:
  pdf_document:
    toc: false
    keep_tex: false
geometry: margin=0.75in
fontsize: 10pt
classoption: letterpaper
# No `title:` field — we render a custom centered title block in the body
# so we get full control over color/font/spacing (and avoid the duplicate
# LaTeX-rendered title that the YAML title produces).
header-includes:
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{colortbl}
  - \usepackage[table]{xcolor}
  - \usepackage{helvet}
  - \renewcommand{\familydefault}{\sfdefault}
  - \usepackage{titlesec}
  - \usepackage{titling}
  - \setlength{\droptitle}{-3em}
  - \definecolor{cohcpurple}{HTML}{602D89}
  - \definecolor{cohcblush}{HTML}{F4EDEC}
  - \definecolor{cohcaccent}{HTML}{AC9E94}
  - \definecolor{cohcdeep}{HTML}{251230}
  # Section headers in COHC purple, slightly smaller than default
  - \titleformat*{\section}{\large\bfseries\color{cohcpurple}}
  - \titleformat*{\subsection}{\normalsize\bfseries\color{cohcpurple}}
  # Tighten section spacing
  - \titlespacing*{\section}{0pt}{1.2em}{0.4em}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE,
                      fig.align = "center", dev = "pdf")
suppressPackageStartupMessages({
  library(knitr)
  library(kableExtra)
  library(ggplot2)
  library(dplyr)
  library(scales)
})

# %||% is a Shiny/rlang infix — provide a local fallback so this template
# renders even when no Shiny context is active.
if (!exists("%||%", mode = "function")) {
  `%||%` <- function(a, b) if (is.null(a)) b else a
}

# Strip the names attribute that vapply leaves on its output — when
# named character vectors land in data.frame(), R promotes the names
# to row.names, which kable then renders as a phantom first column.
# Wrap every prettify-style call with this helper.
.un <- function(x) unname(x)

# Abbreviate US News classification labels so the peer-list column
# isn't 30+ chars wide — that's what was eating the Distance column
# off the right edge of the page.
.abbrev_class <- function(label) {
  short_map <- c(
    "National Liberal Arts Colleges" = "Nat. LAC",
    "National Universities"          = "Nat. Univ.",
    "Regional Universities (North)"   = "Reg. Univ. (N)",
    "Regional Universities (South)"   = "Reg. Univ. (S)",
    "Regional Universities (Midwest)" = "Reg. Univ. (MW)",
    "Regional Universities (West)"    = "Reg. Univ. (W)",
    "Regional Colleges (North)"       = "Reg. Coll. (N)",
    "Regional Colleges (South)"       = "Reg. Coll. (S)",
    "Regional Colleges (Midwest)"     = "Reg. Coll. (MW)",
    "Regional Colleges (West)"        = "Reg. Coll. (W)"
  )
  out <- short_map[label]
  ifelse(is.na(out), label, out)
}

# Pretty-printer for theme weight keys. Uses .theme_label() if it was
# injected into the render envir; otherwise falls back to title-case.
.theme_lbl <- function(keys) {
  if (exists(".theme_label", mode = "function")) .theme_label(keys)
  else tools::toTitleCase(gsub("_", " ", keys))
}

# Inputs from the render envir (see build_peer_pdf in download_helpers.R):
#   peer_result, state
#   .SCHOOLS, .prettify_classification, .prettify_control, .describe_pool_filter
meta  <- peer_result$meta
peers <- peer_result$peers
anchor_uid <- meta$anchor_unitid
anchor_row <- .SCHOOLS[.SCHOOLS$unitid == anchor_uid, , drop = FALSE]
anchor_name <- meta$anchor_name %||% (if (nrow(anchor_row)) anchor_row$instnm else "Anchor")

# Joined frame of anchor + peers used by the composition cards.
peer_meta <- .SCHOOLS[.SCHOOLS$unitid %in% peers$unitid, , drop = FALSE]
```

\begin{flushleft}
{\Huge\bfseries\color{cohcpurple}Peer Search Report}\\[4pt]
{\color{cohcaccent}\small Anchor: \textbf{\color{cohcdeep}`r anchor_name`} ~~|~~ `r nrow(peers)` peers ~~|~~ Generated `r format(Sys.time(), '%Y-%m-%d %H:%M')`}
\end{flushleft}

\vspace{0.5em}
\noindent\textcolor{cohcaccent}{\rule{\linewidth}{0.5pt}}

# Search settings

```{r summary-table}
# Friendly theme-weight string. Internal keys (e.g. "student_body")
# get mapped to display labels ("Student body") via .theme_label() so
# the PDF doesn't expose snake_case implementation detail.
tw <- state$theme_weights %||% list()
tw_lbl <- if (length(tw)) {
  parts <- mapply(function(k, v) sprintf("%s = %.2f", .theme_lbl(k), v),
                   names(tw), unlist(tw), SIMPLIFY = TRUE, USE.NAMES = FALSE)
  paste(parts, collapse = "; ")
} else "(all default)"
vw <- state$variable_weights %||% list()
vw_lbl <- if (length(vw)) {
  paste(sprintf("%s = %.2f", names(vw), unlist(vw)), collapse = "; ")
} else "(none)"

summary_df <- data.frame(
  Field = c("Candidate pool",
            "Pool size after filter",
            "Distance metric",
            "Coverage threshold",
            "Theme weights",
            "Variable overrides"),
  Value = c(.describe_pool_filter(state$candidate_pool),
            as.character(meta$candidate_pool_size %||% ""),
            meta$distance_metric %||% "weighted_euclidean",
            sprintf("%.0f%%", 100 * (meta$coverage_threshold %||% NA)),
            tw_lbl,
            vw_lbl),
  check.names      = FALSE,
  stringsAsFactors = FALSE
)
row.names(summary_df) <- NULL  # belt-and-suspenders against rowname leakage

kable(summary_df, format = "latex", booktabs = TRUE,
      col.names = c("", ""),
      align = c("l", "l"),
      linesep = "",
      row.names = FALSE) |>
  kable_styling(latex_options = c("hold_position"),
                full_width = FALSE,
                position   = "left",
                font_size  = 9) |>
  column_spec(1, bold = TRUE, color = "#251230", width = "1.6in") |>
  column_spec(2, width = "5.0in")
```

# Ranked peer list

```{r peer-table}
# Build the table. Note: .un() strips names from prettify output so
# data.frame() doesn't auto-promote them to row.names (which kable would
# then render as a stray leading column — that was the bug in v1 of
# this template).
disp <- data.frame(
  Rank     = as.character(peers$rank),
  School   = peers$instnm,
  Class    = .un(.abbrev_class(.un(.prettify_classification(
                 peers$usnews_classification %||%
                 rep(NA_character_, nrow(peers)))))),
  `USN`    = ifelse(is.na(peers$usnews_rank %||% rep(NA_real_, nrow(peers))),
                     "", as.character(peers$usnews_rank %||% NA_real_)),
  State    = peers$stabbr %||% rep("", nrow(peers)),
  Distance = sprintf("%.3f", peers$distance),
  check.names      = FALSE,
  stringsAsFactors = FALSE
)

if (nrow(anchor_row) == 1) {
  a_row <- data.frame(
    Rank     = "★",
    School   = paste0(anchor_row$instnm, " (anchor)"),
    Class    = .un(.abbrev_class(.un(.prettify_classification(
                   anchor_row$usnews_classification)))),
    `USN`    = if (is.na(anchor_row$usnews_rank)) "" else
               as.character(anchor_row$usnews_rank),
    State    = anchor_row$stabbr %||% "",
    Distance = "0.000",
    check.names      = FALSE,
    stringsAsFactors = FALSE
  )
  disp <- rbind(a_row, disp)
}
row.names(disp) <- NULL

n_anchor <- if (nrow(anchor_row) == 1) 1 else 0

kt <- kable(disp, format = "latex", booktabs = TRUE, longtable = TRUE,
            align = c("r", "l", "l", "r", "c", "r"),
            linesep = "",
            row.names = FALSE) |>
  # No "striped" here — we shade the anchor row in cohcblush below, and
  # an alternating stripe in the same color would hide it.
  kable_styling(latex_options = c("repeat_header"),
                font_size     = 9,
                position      = "left") |>
  column_spec(1, width = "0.5in") |>
  column_spec(2, width = "2.6in") |>
  column_spec(3, width = "1.4in") |>
  column_spec(4, width = "0.5in") |>
  column_spec(5, width = "0.4in") |>
  column_spec(6, width = "0.6in") |>
  row_spec(0, bold = TRUE, color = "#602D89")

# Highlight the anchor row in COHC purple.
if (n_anchor == 1) {
  kt <- kt |> row_spec(1, bold = TRUE,
                       background = "F4EDEC",
                       color      = "#251230")
}

kt
```

# Composition of the peer set

```{r composition-cards, fig.width=6.8, fig.height=6, fig.pos="H"}
# Same 5 dimensions the on-screen Composition tab uses. Replicated
# here so the Rmd is self-contained.
.col_or <- function(df, name) {
  if (name %in% names(df)) df[[name]] else rep(NA_character_, nrow(df))
}

state_to_region <- function(stabbr) {
  northeast <- c("CT","ME","MA","NH","NJ","NY","PA","RI","VT")
  midwest   <- c("IL","IN","IA","KS","MI","MN","MO","NE","ND","OH","SD","WI")
  south     <- c("AL","AR","DE","DC","FL","GA","KY","LA","MD","MS","NC",
                 "OK","SC","TN","TX","VA","WV")
  west      <- c("AK","AZ","CA","CO","HI","ID","MT","NV","NM","OR","UT","WA","WY")
  out <- rep(NA_character_, length(stabbr))
  out[stabbr %in% northeast] <- "Northeast"
  out[stabbr %in% midwest]   <- "Midwest"
  out[stabbr %in% south]     <- "South"
  out[stabbr %in% west]      <- "West"
  out
}

dims <- list(
  list(label = "US News classification",
       vals  = .un(.abbrev_class(.un(.prettify_classification(
                   .col_or(peer_meta, "usnews_classification"))))),
       anchor_val = if (nrow(anchor_row))
                      .un(.abbrev_class(.un(.prettify_classification(
                          anchor_row$usnews_classification))))
                    else NA),
  list(label = "Institution size (Carnegie)",
       vals  = as.character(.col_or(peer_meta, "ic2025size_label")),
       anchor_val = if (nrow(anchor_row))
                      as.character(anchor_row$ic2025size_label %||% NA)
                    else NA),
  list(label = "Geographic region",
       vals  = state_to_region(.col_or(peer_meta, "stabbr")),
       anchor_val = if (nrow(anchor_row)) state_to_region(anchor_row$stabbr)
                    else NA),
  list(label = "Carnegie Research Activity",
       vals  = as.character(.col_or(peer_meta, "research2025_label")),
       anchor_val = if (nrow(anchor_row))
                      as.character(anchor_row$research2025_label %||% NA)
                    else NA),
  list(label = "Athletics division",
       vals  = ifelse(is.na(.col_or(peer_meta, "athletics_division")),
                       "(no athletics)",
                       as.character(.col_or(peer_meta, "athletics_division"))),
       anchor_val = if (nrow(anchor_row))
                      ifelse(is.na(anchor_row$athletics_division %||% NA),
                             "(no athletics)",
                             as.character(anchor_row$athletics_division %||% NA))
                    else NA)
)

card_df <- do.call(rbind, lapply(dims, function(d) {
  tab <- table(d$vals, useNA = "no")
  if (length(tab) == 0) return(NULL)
  pct <- as.numeric(tab) / sum(tab) * 100
  is_anchor <- names(tab) == d$anchor_val
  data.frame(
    dimension = d$label,
    category  = names(tab),
    n         = as.integer(tab),
    pct       = pct,
    is_anchor = is_anchor,
    stringsAsFactors = FALSE
  )
}))

card_df$category_lbl <- ifelse(card_df$is_anchor,
                                paste0("★ ", card_df$category),
                                card_df$category)
# Truncate very long labels so bars stay readable inside the half-page card.
card_df$category_lbl <- ifelse(nchar(card_df$category_lbl) > 32,
                                paste0(substr(card_df$category_lbl, 1, 29),
                                       "..."),
                                card_df$category_lbl)

# Sort categories within each card by share descending.
card_df <- card_df[order(card_df$dimension, -card_df$pct), ]
card_df$category_lbl <- factor(card_df$category_lbl,
                                levels = rev(unique(card_df$category_lbl)))

ggplot(card_df, aes(x = pct, y = category_lbl, fill = is_anchor)) +
  geom_col(width = 0.7) +
  geom_text(aes(label = sprintf("%d (%d%%)", n, round(pct))),
            hjust = -0.08, size = 2.3, color = "#251230") +
  facet_wrap(~ dimension, scales = "free_y", ncol = 2) +
  # Bumped right expansion from 18% to 40% so the "57 (78%)" labels
  # render in full instead of clipping mid-string.
  scale_x_continuous(expand = expansion(mult = c(0, 0.40)), labels = NULL) +
  scale_fill_manual(values = c("TRUE" = "#602D89", "FALSE" = "#AC9E94"),
                    guide = "none") +
  labs(x = NULL, y = NULL) +
  theme_minimal(base_size = 8) +
  theme(
    text       = element_text(family = "Helvetica"),
    strip.text = element_text(face = "bold", color = "#602D89",
                              size = 9, hjust = 0),
    panel.grid.major.x = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.y = element_blank(),
    axis.ticks         = element_blank(),
    plot.margin        = margin(2, 8, 2, 2)
  )
```

\vspace{0.5em}
\noindent\textcolor{cohcaccent}{\rule{\linewidth}{0.4pt}}

\begin{center}
{\color{cohcaccent}\footnotesize Generated by the Peer Schools Explorer. For methodology details see the Help tab in the app.}
\end{center}
