---
title: "Peer Search Report"
output:
  pdf_document:
    toc: false
    keep_tex: false
geometry: margin=0.75in
fontsize: 10pt
# mainfont only works with xelatex/lualatex; we stick with the default
# pdflatex for portability. Letter paper for the US institutional
# research audience.
classoption: letterpaper
header-includes:
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage{wrapfig}
  - \usepackage{float}
  - \usepackage{colortbl}
  - \usepackage[table]{xcolor}
  - \definecolor{cohcpurple}{HTML}{602D89}
  - \definecolor{cohcblush}{HTML}{F4EDEC}
  - \definecolor{cohcaccent}{HTML}{AC9E94}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE,
                      fig.align = "center", dev = "pdf")
suppressPackageStartupMessages({
  library(knitr)
  library(ggplot2)
  library(dplyr)
  library(scales)
})

# kableExtra is optional — fall back to plain kable when absent so the
# render doesn't hard-error on hosts without it.
kableExtra_or_plain <- function(x) {
  if (requireNamespace("kableExtra", quietly = TRUE)) {
    kableExtra::kable_styling(x,
      latex_options = c("hold_position", "striped"),
      stripe_color = "cohcblush",
      full_width   = FALSE,
      position     = "left",
      font_size    = 9)
  } else {
    x
  }
}

# %||% is a Shiny/rlang infix that returns the LHS unless it's NULL.
# Define a local fallback so this template renders even when no Shiny
# context is active.
if (!exists("%||%", mode = "function")) {
  `%||%` <- function(a, b) if (is.null(a)) b else a
}

# The render envir has these injected:
#   peer_result, state
#   .SCHOOLS, .prettify_classification, .prettify_control, .describe_pool_filter
# (See build_peer_pdf in shiny_app/R/download_helpers.R.)
meta  <- peer_result$meta
peers <- peer_result$peers
anchor_uid <- meta$anchor_unitid
anchor_row <- .SCHOOLS[.SCHOOLS$unitid == anchor_uid, , drop = FALSE]
anchor_name <- meta$anchor_name %||% (if (nrow(anchor_row)) anchor_row$instnm else "Anchor")

# Joined frame of anchor + peers used by the composition cards.
peer_meta <- .SCHOOLS[.SCHOOLS$unitid %in% peers$unitid, , drop = FALSE]
```

\begin{center}
{\color{cohcpurple}\Large\bfseries Peer Search Report}\\[2pt]
{\color{cohcaccent}\small Anchor: `r anchor_name` ~~|~~ `r nrow(peers)` peers ~~|~~ Generated `r format(Sys.time(), '%Y-%m-%d %H:%M')`}
\end{center}

\vspace{-0.5em}

## Search settings

```{r summary-table, results="asis"}
tw <- state$theme_weights %||% list()
tw_lbl <- if (length(tw))
  paste(sprintf("%s = %.2f", names(tw), unlist(tw)), collapse = "; ") else "(all default)"
vw <- state$variable_weights %||% list()
vw_lbl <- if (length(vw))
  paste(sprintf("%s = %.2f", names(vw), unlist(vw)), collapse = "; ") else "(none)"

summary_df <- data.frame(
  Field = c("Candidate pool",
            "Pool size after filter",
            "Distance metric",
            "Coverage threshold",
            "Theme weights",
            "Variable overrides"),
  Value = c(.describe_pool_filter(state$candidate_pool),
            as.character(meta$candidate_pool_size %||% ""),
            meta$distance_metric %||% "weighted_euclidean",
            sprintf("%.0f%%", 100 * (meta$coverage_threshold %||% NA)),
            tw_lbl,
            vw_lbl),
  check.names = FALSE,
  stringsAsFactors = FALSE
)
kable(summary_df, format = "latex", booktabs = TRUE,
      col.names = c("", ""),
      align = c("l", "l"), linesep = "") |>
  kableExtra_or_plain()
```

## Ranked peer list

```{r peer-table, results="asis"}
# Anchor row at top, then peers in rank order. Distances formatted to 3 dp.
disp <- data.frame(
  Rank   = peers$rank,
  School = peers$instnm,
  Class  = .prettify_classification(peers$usnews_classification %||%
                                      rep(NA_character_, nrow(peers))),
  `USN Rank` = ifelse(is.na(peers$usnews_rank %||% rep(NA_real_, nrow(peers))),
                       "", as.character(peers$usnews_rank %||% NA_real_)),
  State    = peers$stabbr %||% rep("", nrow(peers)),
  Distance = sprintf("%.3f", peers$distance),
  check.names = FALSE,
  stringsAsFactors = FALSE
)

if (nrow(anchor_row) == 1) {
  a_row <- data.frame(
    Rank   = "*",
    School = paste0(anchor_row$instnm, " (anchor)"),
    Class  = .prettify_classification(anchor_row$usnews_classification),
    `USN Rank` = if (is.na(anchor_row$usnews_rank)) "" else
                 as.character(anchor_row$usnews_rank),
    State    = anchor_row$stabbr,
    Distance = "0.000",
    check.names = FALSE,
    stringsAsFactors = FALSE
  )
  disp <- rbind(a_row, disp)
}

kable(disp, format = "latex", booktabs = TRUE, longtable = TRUE,
      align = c("r", "l", "l", "r", "c", "r"),
      linesep = "") |>
  kableExtra_or_plain()
```

## Composition of the peer set

```{r composition-cards, fig.width=6.8, fig.height=6, fig.pos="H"}
# Reuse the same 5 dimensions the Composition tab shows. Replicated
# here so the Rmd is self-contained — when these definitions change
# in mod_peer_table.R, mirror them here.
.col_or <- function(df, name) {
  if (name %in% names(df)) df[[name]] else rep(NA_character_, nrow(df))
}

# Geographic region from state — mirrors .REGIONS in ui_sidebar.R but
# we only need the top-level four buckets here.
state_to_region <- function(stabbr) {
  northeast <- c("CT","ME","MA","NH","NJ","NY","PA","RI","VT")
  midwest   <- c("IL","IN","IA","KS","MI","MN","MO","NE","ND","OH","SD","WI")
  south     <- c("AL","AR","DE","DC","FL","GA","KY","LA","MD","MS","NC",
                 "OK","SC","TN","TX","VA","WV")
  west      <- c("AK","AZ","CA","CO","HI","ID","MT","NV","NM","OR","UT","WA","WY")
  out <- rep(NA_character_, length(stabbr))
  out[stabbr %in% northeast] <- "Northeast"
  out[stabbr %in% midwest]   <- "Midwest"
  out[stabbr %in% south]     <- "South"
  out[stabbr %in% west]      <- "West"
  out
}

# Build a long frame: one row per (dimension, peer school, category value).
dims <- list(
  list(label = "US News classification",
       vals  = .prettify_classification(.col_or(peer_meta, "usnews_classification")),
       anchor_val = if (nrow(anchor_row)) .prettify_classification(anchor_row$usnews_classification) else NA),
  list(label = "Institution size (Carnegie)",
       vals  = as.character(.col_or(peer_meta, "ic2025size_label")),
       anchor_val = if (nrow(anchor_row)) as.character(anchor_row$ic2025size_label %||% NA) else NA),
  list(label = "Geographic region",
       vals  = state_to_region(.col_or(peer_meta, "stabbr")),
       anchor_val = if (nrow(anchor_row)) state_to_region(anchor_row$stabbr) else NA),
  list(label = "Carnegie Research Activity",
       vals  = as.character(.col_or(peer_meta, "research2025_label")),
       anchor_val = if (nrow(anchor_row)) as.character(anchor_row$research2025_label %||% NA) else NA),
  list(label = "Athletics division",
       vals  = ifelse(is.na(.col_or(peer_meta, "athletics_division")),
                       "(no athletics)",
                       as.character(.col_or(peer_meta, "athletics_division"))),
       anchor_val = if (nrow(anchor_row))
                      ifelse(is.na(anchor_row$athletics_division %||% NA),
                             "(no athletics)",
                             as.character(anchor_row$athletics_division %||% NA))
                    else NA)
)

card_df <- do.call(rbind, lapply(dims, function(d) {
  tab <- table(d$vals, useNA = "no")
  if (length(tab) == 0) return(NULL)
  pct <- as.numeric(tab) / sum(tab) * 100
  is_anchor <- names(tab) == d$anchor_val
  data.frame(
    dimension = d$label,
    category  = names(tab),
    n         = as.integer(tab),
    pct       = pct,
    is_anchor = is_anchor,
    stringsAsFactors = FALSE
  )
}))

# Star the anchor category in the bar label.
card_df$category_lbl <- ifelse(card_df$is_anchor,
                                paste0("★ ", card_df$category),
                                card_df$category)
# Truncate very long labels so the bars stay readable.
card_df$category_lbl <- ifelse(nchar(card_df$category_lbl) > 40,
                                paste0(substr(card_df$category_lbl, 1, 37), "..."),
                                card_df$category_lbl)

# Sort categories within each card by share descending so the longest
# bar is on top of each subplot.
card_df <- card_df[order(card_df$dimension, -card_df$pct), ]
card_df$category_lbl <- factor(card_df$category_lbl,
                                levels = rev(unique(card_df$category_lbl)))

ggplot(card_df, aes(x = pct, y = category_lbl, fill = is_anchor)) +
  geom_col(width = 0.7) +
  geom_text(aes(label = sprintf("%d (%d%%)", n, round(pct))),
            hjust = -0.05, size = 2.6, color = "#251230") +
  facet_wrap(~ dimension, scales = "free_y", ncol = 2) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.18)), labels = NULL) +
  scale_fill_manual(values = c("TRUE" = "#602D89", "FALSE" = "#AC9E94"),
                    guide = "none") +
  labs(x = NULL, y = NULL) +
  theme_minimal(base_size = 8) +
  theme(
    strip.text = element_text(face = "bold", color = "#602D89",
                              size = 9, hjust = 0),
    panel.grid.major.x = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.y = element_blank(),
    axis.ticks         = element_blank(),
    plot.margin = margin(2, 6, 2, 2)
  )
```

\vfill
\begin{center}
{\color{cohcaccent}\small Generated by the Peer Schools Explorer. \\ For methodology details see the Help tab in the app.}
\end{center}
